A High-Level Embedded Language for Low-Level GPU Kernel Programming

### 2013-11-04 src2/hl/paperThesis.tex
update

Obsidian is an embedded language for GPU programming that aims to raise the 
level of abstraction for the GPU programmer, while maintaining control over 
the low-level details that are needed for performance. 
So far,
we have chosen not to try to apply automatic transformations or compiler 
optimisation techniques to a GPU architecture agnostic program and hope 
to get performance. In Obsidian, we want to put the tools to make the 
correct decisions into the hands of the programmer. From Obsidian programs, 
CUDA kernels are generated, while host code must still be written in CUDA.
For key building blocks like scan and reduce, we see a trend towards
the use of fewer and much more complicated kernels~\citehl{merrill}, so
that the host code is often relatively simple. We wish to support the construction of these sophisticated kernels.

@@

The last step in compilation process uses the Language.C.Quote library to 
turn the {\tt IM} into CUDA code. This compilation stage takes a {\em configuration} 
as parameter. The configuration tells how many actual threads per block the 
CUDA code should be generated for. At this stage parallel for loops (\tt SForAll) 
are turned into more than one parallel for loop in sequence, if needed.

@@

In reference ~\citehl{CSORT}, we experiment with atomic operations 
and are forced to write our code in a very imperative style. The addition of mutable 
arrays and a set of operations on them could improve this situation 
while giving up purity. We will explore what the addition of mutable arrays to 
Obsidian would mean as future work.

>>>>

Obsidian is an embedded language for GPU programming that aims to raise the 
level of abstraction for the GPU programmer, while maintaining control over 
the low-level details that are needed for performance.

So far, we have chosen not to try to apply automatic transformations or compiler 
optimisation techniques to a GPU architecture agnostic program and hope 
to get performance. In Obsidian, we want to put the tools to make the 
correct decisions into the hands of the programmer. From Obsidian programs, 
CUDA kernels are generated; host code can be written in Obsidian, but is still 
preferably written in CUDA.

For key building blocks like scan and reduce, we see a trend towards
the use of fewer and much more complicated kernels~\citehl{merrill}, so
that the host code is often relatively simple. We wish to support the construction of these sophisticated kernels.

@@

The last step in compilation process uses the Language.C.Quote library to 
turn the {\tt IM} into CUDA code. This compilation stage takes a {\em configuration} 
as parameter. The configuration tells how many actual threads per block the 
CUDA code should be generated for. At this stage parallel for loops ({\tt SForAll})
are turned into more than one parallel for loop in sequence, if needed.

@@

In reference ~\citehl{JOSEFJOELCSORT}, we experiment with atomic operations 
and are forced to write our code in a very imperative style. The addition of mutable 
arrays and a set of operations on them could improve this situation 
while giving up purity. We will explore what the addition of mutable arrays to 
Obsidian would mean as future work.
